---
title: "WUP_04"
author: "Erick Barraza!"
date: "Summer 2021"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

```{r}
gssdata=read_csv("GSScsv.csv")
```



# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(gssdata,aes(x=DEGREE,y=WORDSUM))
```
## Now add in boxplot

```{r}
basicCN + geom_boxplot()
```

  
The boxplot above shows the difference between WORDSUM and degrees. If we were to move the LT High School over to the right, it would not be in the of the Junior College boxplot. This would also apply to the Graduate boxplot which is different from the High School boxplot and LT High school plot if they were moved next to each other. They would not interfere with each other which means that they appear to be different.  

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=gssdata)
```

The numerical results show that the highest mean is the graduate degree and the lowest being LT High School degree being 3 points away from it at 7.750000 which makes it the smallest. The Graduate degree had the highest mean at 10.095238 while the closest where junior college at 9.916667 and bachelor at 9.204545 which means that there is a difference between all five degrees mean.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=gssdata)
summary.aov(model1)
```

The p-value here is very small at 0.00236 which means that there is a small chance of this occuring and that we could reject the null hypothesis. 

```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 13.2, 2, 2))
plot(TukeyHSD(model1),las=2)
```

Looking at this graph shows that there are differences between bachelor and LT High School when it comes to their WORDSUM. We know this since the line for 0 does not cross LT High School and High School graduate which means they are very different from a bachelor degree.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Looking at what null hypothesis stating that there is no difference between the WORDSUM of a bachelor degree and high school degree. First we added a boxplot in order to give a visual description that would show how different certain boxplots are from another. From there it showed that if we stacked the bachelors boxplot and LT High School boxplot, they would not interefere with each other. Then we added a numerical results that showed that the highest mean was the bachelors degree and the lowest was LT High School which was almost 3 points lower. After that I added a summary for the p-value which was 0.000236 which was a very low number and meant that we could reject the null hypothesis. Next was the Tukey model which showed how much of a difference each degree had within each other and their WORDSUM. Once this was done it showed that there are difference between bachelor, high school graduate, and LT high school. This meant that we can use the alternate hypothesis since we have data to back it up  